---
title: "Beers and Breweries Analysis"
author: "Joel Laskow and Renu Karthikeyan"
date: "2023-10-21"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

#Statement of Purpose:
Esteemed members of Budweiser's leadership team, today, we have the privilege of presenting a comprehensive analysis of beers and brewery data, aimed at uncovering key insights and trends that can significantly impact our business strategy.The purpose of this document is to address lingering questions regarding beer production within the Budweiser brewing company, and to make this information accessible and actionable. 

The following presentation highlights both the exciting discoveries we have made and the actionable insights derived from the data. Thank you for your time, and we look forward to empowering Budweiser with data-driven wisdom.


```{r}
# Libraries Used
library(base)
library(dplyr)
library(caret)
library(tidyverse)
library(ggplot2)
library(naniar)
library(class)
library(GGally)
```

```{r}
#Breweries <- read.csv("/cloud/project/DS 6306/Datasets/Breweries.csv")
#Beers <- read.csv("/cloud/project/DS 6306/Datasets/Beers (1).csv")


setwd("/Users/coolp/Desktop/Learning/SMU/Fall 2023/DS 6306/group project 1")

breweries <- read.csv("/Users/coolp/Desktop/Learning/SMU/Fall 2023/DS 6306/group project 1/Breweries.csv")
beers <- read.csv("/Users/coolp/Desktop/Learning/SMU/Fall 2023/DS 6306/group project 1/Beers.csv")

#Brewery Data
breweries<-data.frame(breweries)

# Beer Data 
beers<-data.frame(beers)
colnames(beers)[5]="Brew_ID"

```


```{r}

# Merged Data Sets by brewery ID:

df_bud<-merge(beers,breweries, by="Brew_ID")

```



# First 6 Rows of the Merged Dataset
```{r}

head(df_bud, 6)

```

# Final 6 Rows of the Merged Dataset
```{r}
tail(na.omit(df_bud), 6)

```



# Addressing NA Values

```{r}
# We visualize the position of all NA values within the dataset

vis_miss(df_bud) 
```
```{r}
summary(df_bud)

```
 - Visualization of all missing (NA) values indicates 3% of ABV Values are reported as NA. 42% of IBU values (1005) are reported as NA. 
 - Removing this quantity of missing values would substantially reduced the size of our dataset
 - Solution:
   The missing values show no apparent pattern in the data set, suggesting that they are randomly distributed. Under this condition, it is beneficial to impute the missing values with the respective mean ABV and IBU values.


```{r}
#Impute NA Values with mean from IBU and ABV columns

df_bud$IBU[is.na(df_bud$IBU)] <- mean(df_bud$IBU, na.rm = TRUE)
df_bud$ABV[is.na(df_bud$ABV)] <- mean(df_bud$ABV, na.rm = TRUE)

```


 
# How many breweries are in each states?
```{r}
breweries$State=factor(breweries$State)



breweries_per_state <- breweries %>% group_by(State)%>%tally(name="count")
 
breweries_per_state

ggplot(breweries_per_state, aes(x = State, y = count, fill = State)) +
  geom_bar(stat="identity", width=0.7) +
  theme(axis.text.x=element_text(angle=45, hjust=1))

 
```
-The states with the most breweries are Colorado and California. Economic fluctuations within these 2 states might have the greatest impact on production (like droughts, recessions, or supply chain issues)


# Calculate median alcohol content by State:

```{r}
#Build a new dataframe for median alcohol content by state

ABV_Summary_by_State<-aggregate(df_bud$ABV, list(df_bud$State),median)

ABV<-data.frame(ABV_Summary_by_State)

names(ABV)[1]<-"State"
names(ABV)[2]<-"Median_ABV"



# Build a new dataframe for median IBU by state
IBU_Summary_by_State<-aggregate(df_bud$IBU,list(df_bud$State),median)
  
IBU<-data.frame(IBU_Summary_by_State)

names(IBU)[1]<-"State"
names(IBU)[2]<-"Median_IBU"
 
```


# Median IBU By State

```{r}

ggplot(IBU, aes(x=State, y=Median_IBU))+
  geom_bar(stat="identity", position="dodge", color="black")+
  labs(title="Median Alcohol Content (IBU) By State", xlab="State",ylab="Median_ABV")+
  theme_minimal()

```

The state with the highest IBU (most bitterness) is West Virginia. Here, we see that the means are all homogenuous, and it is in line with the imputation done earlier. 

#Median ABV by State

```{r}
ABV_Plot <- ggplot(ABV, aes(x = factor(State), y = Median_ABV)) +
  geom_bar(stat="identity", width=0.9, color="black", position = position_dodge(width = 0.9)) +
  labs(title = "Median ABV by State", x="State", y="Alcohol by Volume (ABV)") +
  theme_minimal(axis.text.x = element_text(angle = 75, hjust = 1)) 

print(ABV_Plot)



```

The states with highest alcohol content are 1. District of Columbia, 2. Kentucky, 3.Michigan, 4. New Mexico, and 5. West Virginia


```{r}

library(ggplot2)

IBU_Plot <- ggplot(IBU, aes(x = State, y = Median_IBU)) +
  geom_bar(stat = "identity", width = 0.7, fill = "blue") +  # Adjusted width and added fill color
  labs(title = "Median IBU by State") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

IBU_Plot



```


# State With Maximum ABV
```{r}
max_abv_row <- df_bud[which.max(df_bud$ABV), ]

# Print the state with the maximum ABV
print("State with Maximum ABV:")
print(max_abv_row$State)
```

#State With Max Median ABV
```{r}

ABV[ABV$Median_ABV==max(ABV$Median_ABV),]

```

# State with bitterest beer (by IBU)

```{r}

max_IBU_row <- df_bud[which.max(df_bud$IBU), ]

# Print the state with the maximum ABV
print("State with Maximum IBU:")
print(max_IBU_row$State) 

```
 


# State with highest Median Bitterness
```{r}
IBU[IBU$Median_IBU==max(IBU$Median_IBU),] 
```


# ABV Distribution:

```{r}
df_bud %>%
  ggplot(aes(x = ABV)) +
  geom_histogram(binwidth = 0.001, fill = "purple", color = "black", alpha = 0.5) +
  geom_density(color = "green") +  
  labs(title = "Distribution of Alcohol by Volume", xlab="ABV", ylab="Count")


```


#Observations on Bitterness vs ABV:
```{r}


ggplot(df_bud, aes(x=ABV, y=IBU, na.rm=TRUE))+
  geom_point(aes(color=factor(Style)))+
  stat_smooth(method="lm",linewidth=1)

ggplot(df_bud, aes(x=IBU, y=ABV, na.rm=TRUE))+
  geom_point(aes(color=factor(Style)))+
  stat_smooth(method="lm",size=1)+
  labs(title="IBU vs ABV")


```



 

# We want to use IBU and ABV to predict whether or not a beer is an Ale

```{r}

# We will first filter out all styles that are not Ales 

df_Ales <- df_bud %>% filter(grepl("Ale|IPA|PA", Style, ignore.case = TRUE))


#We will classify out filtered beers by "IPA" or "Not_IPA"
df_Ales$Is_IPA <- ifelse(grepl("Indian Pale Ale|IPA", df_Ales$Style, ignore.case = TRUE), "IPA", "Non_IPA")

df_Ales$Is_IPA<-factor(df_Ales$Is_IPA) 
  
```


```{r}

iterations <- 500
numks <- 10
percentsplit <- 0.8

# Initialize an empty data frame to store results
results_df <- data.frame(accuracy = numeric(0), k = numeric(0))

for (j in 1:iterations) {
  trainIndices <- sample(1:nrow(df_Ales), round(percentsplit * nrow(df_Ales)))
  train <- df_Ales[trainIndices, ]
  test <- df_Ales[-trainIndices, ]
  
  for (i in 1:numks) {
    classifications <- knn(train[, c("ABV", "IBU")], test[, c("ABV", "IBU")], train$Is_IPA, prob = TRUE, k = i)
    CM <- confusionMatrix(table(classifications, test$Is_IPA))
    
    # Append the accuracy and k value to the results data frame
    results_df <- rbind(results_df, data.frame(accuracy = CM$overall["Accuracy"], k = i))
  } 
}

results_df<-aggregate(results_df$accuracy,list(results_df$k),mean)

results_df_named<-results_df
names(results_df_named) <- c("k", "Average_Accuracy")

```

```{r}

#We will build an Acc vs K plot to assess optimal k

ggplot(results_df_named, aes(x = k, y = `Average_Accuracy`)) +
  geom_line()+
  labs(title = "Average Accuracy vs K", xlab="k", ylab="Average Accuracy")

```

# Accuracy plot shows 5 as optimal k:

```{r}


trainIndex <- sample(1:nrow(df_Ales), round(percentsplit * nrow(df_Ales)))
  train <- df_Ales[trainIndex, ]
  test <- df_Ales[-trainIndex, ]
  
    classifications <- knn(train[, c("ABV", "IBU")], test[, c("ABV", "IBU")], train$Is_IPA, prob = TRUE, k = 5)
    CM <- confusionMatrix(table(classifications, test$Is_IPA))
 

CM
``` 
Final knn model delivers 83.86% accuracy. Accuracy may differ with higher iterations during training. 




# We next sought to gather some extra insights from our datasets
```{r}

# Renu charted ABV associated with different styles of beer
df_bud %>% ggplot(aes(x = ABV, y = State, color = Ounces)) + geom_point(alpha = 0.7) + 
  labs(title = "ABV by Style", x = "ABV", y = "Style")
```

```{r}

# Renu then charted the average volume in Ounces per beer by State. The results suggest Colorado (CO) produces the largest beers
df_bud %>% ggplot(aes(x = State, color = Ounces)) + geom_bar(alpha = 0.7) + 
  labs(title = "Ounces by State", x = "State", y = "Ounces") + theme_minimal()
```

```{r}

# Renu then sought to compare the mean ounces associated with each style of beer
df_bud %>%

  ggplot(aes(x = reorder(Style, -Ounces), y = Ounces)) +

  geom_bar(stat = "identity", alpha = 0.7) +

  labs(title = "Ounces by Style", x = "Style", y = "Ounces") +

  theme_minimal() +

  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```





```{r}
# Joel built a dataframe of the % ABV misreports per State
eda_bud<-merge(beers,breweries, by="Brew_ID")

eda_bud<-eda_bud%>%mutate(ABV_Errors=ifelse(is.na(ABV), 1, 0))



eda_bud<-eda_bud%>%mutate(IBU_Errors=ifelse(is.na(IBU), 1,0))




agg_ABV_NA<- aggregate(eda_bud$ABV_Errors, by=list(eda_bud$State), FUN=sum)


colnames(agg_ABV_NA)<-c("State","Brews_With_Missing_ABV")

totalbeers<-(aggregate(eda_bud$ABV_Errors,by=list(eda_bud$State), FUN=length))

colnames(totalbeers)<-c("State","Total_Brews_From_State")


agg_ABV_NA<-data.frame(agg_ABV_NA)



agg_ABV_NA$Percent<-agg_ABV_NA$Brews_With_Missing_ABV/totalbeers$Total_Brews_From_State



ABV_NAs_Per_State<-merge(agg_ABV_NA,totalbeers, by="State")

```
```{r}
# Joel then built a scatterplot comparing variety of beers produced to percent of ABVs misreported

ggplot(ABV_NAs_Per_State, aes(x = Total_Brews_From_State, y = Percent)) +
  geom_point() +
  geom_smooth(method=lm)+
  labs(x = "Brews Produced", y = "Percentage of Brews With Missing ABVs")
```

# IBU Misreport was Compared to Beer Variety by State
```{r}

# Joel built a dataframe of the % ABV misreports per State
eda_bud<-merge(beers,breweries, by="Brew_ID")

eda_bud<-eda_bud%>%mutate(ABV_Errors=ifelse(is.na(ABV), 1, 0))



eda_bud<-eda_bud%>%mutate(IBU_Errors=ifelse(is.na(IBU), 1,0))




agg_IBU_NA<- aggregate(eda_bud$IBU_Errors, by=list(eda_bud$State), FUN=sum)


colnames(agg_IBU_NA)<-c("State","Brews_With_Missing_IBU")

totalbeers<-(aggregate(eda_bud$IBU_Errors,by=list(eda_bud$State), FUN=length))

colnames(totalbeers)<-c("State","Total_Brews_From_State")


agg_IBU_NA<-data.frame(agg_IBU_NA)



agg_IBU_NA$Percent<-agg_IBU_NA$Brews_With_Missing_IBU/totalbeers$Total_Brews_From_State



IBU_NAs_Per_State<-merge(agg_IBU_NA,totalbeers, by="State")

head(IBU_NAs_Per_State)

 
ggplot(IBU_NAs_Per_State, aes(x = Total_Brews_From_State, y = Percent)) +
  geom_point() +
  geom_smooth(method=lm)+
  labs(title="IBU Misreport Frequency vs Beer Variety", x = "Brews Produced", y = "Percentage of Brews With Missing IBUs")
```